Testing gene methylation enrichment analysis approaches using simulated data

### 2023-10-18 figs/GSE158422_simulate.Rmd
summarising simulation data

# enrich
ttenrich <- function(m,genesets,cores=1) {
  res <- mclapply( 1:length(genesets), function(i) {
    gs <- genesets[i]
    name <- names(gs)
    n_members <- length(which(rownames(m) %in% gs[[1]]))
    if ( n_members > 4 ) {
      tstats <- m[which(rownames(m) %in% gs[[1]]),]
      myn <- length(tstats)
      mymean <- mean(tstats)
      mymedian <- median(tstats)
      wt <- t.test(tstats)
      res <- c(name,myn,mymean,mymedian,wt$p.value)
    }
  } , mc.cores = cores)
  res_df <- do.call(rbind, res)
  rownames(res_df) <- res_df[,1]
  res_df <- res_df[,-1]
  colnames(res_df) <- c("n_genes","t_mean","t_median","pval")
  tmp <- apply(res_df,2,as.numeric)
  rownames(tmp) <- rownames(res_df)
  res_df <- tmp
  res_df <- as.data.frame(res_df)
  res_df <- res_df[order(res_df$pval),]
  res_df$logp <- -log10(res_df$pval )
  res_df$fdr <- p.adjust(res_df$pval,method="fdr")
  res_df[order(abs(res_df$pval)),]
  return(res_df)
}

simla <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  dm3 <- runlimma(mval=mval2,design=d,myann=myann)
  dmagg1 <- agg(dm3,cores=4)
  m1 <- dmagg1$gme_res_df[,"mean",drop=FALSE]
  lares1 <- ttenrich(m=m1,genesets=gsets,cores=4)
  gsig_up3 <- rownames(subset(lares1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(lares1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(lares1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

@@

simal <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  # al pipeline
  medf1 <- chragg(mval=mval2,myann=myann,cores=4)
  magg1 <- agglimma(medf1,d)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  alres1 <- ttenrich(m=m1,genesets=gsets,cores=4)
  # summarise results
  gsig_up3 <- rownames(subset(alres1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(alres1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(alres1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

@@

gres <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- lapply(1:sims,function(i) {
    simgsa(genesetdatabase=gs_symbols, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5,
      frac_probes=0.5, groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  })
  res <- do.call(rbind,res)
  return(res)
})

@@

@@

@@

lares <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:sims,function(i) {
    simla(genesetdatabase=gs_symbols, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,res)
  return(res)
})

lares

@@

alres <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:50,function(i) {
    simal(genesetdatabase=gs_symbols, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,alres)
  return(res)
})

@@

>>>>

# enrich parametric
ttenrich <- function(m,genesets,cores=1,testtype="selfcontained") {
  res <- mclapply( 1:length(genesets), function(i) {
    scores <- m[,1]
    gs <- genesets[i]
    name <- names(gs)
    n_members <- length(which(rownames(m) %in% gs[[1]]))
    if ( n_members > 4 ) {
      tstats <- m[which(rownames(m) %in% gs[[1]]),]
      myn <- length(tstats)
      mymean <- mean(tstats)
      mymedian <- median(tstats)
      if ( testtype == "selfcontained" ) { wt <- t.test(tstats) }
      if ( testtype == "competitive" ) { wt <- t.test(tstats,scores) }
      res <- c(name,myn,mymean,mymedian,wt$p.value)
    }
  } , mc.cores = cores)
  res_df <- do.call(rbind, res)
  rownames(res_df) <- res_df[,1]
  res_df <- res_df[,-1]
  colnames(res_df) <- c("n_genes","t_mean","t_median","pval")
  tmp <- apply(res_df,2,as.numeric)
  rownames(tmp) <- rownames(res_df)
  res_df <- tmp
  res_df <- as.data.frame(res_df)
  res_df <- res_df[order(res_df$pval),]
  res_df$logp <- -log10(res_df$pval )
  res_df$fdr <- p.adjust(res_df$pval,method="fdr")
  res_df[order(abs(res_df$pval)),]
  return(res_df)
}

# enrich non-parametric
wtenrich <- function(m,genesets,cores=1,testtype="selfcontained") {
  res <- mclapply( 1:length(genesets), function(i) {
    scores <- m[,1]
    gs <- genesets[i]
    name <- names(gs)
    n_members <- length(which(rownames(m) %in% gs[[1]]))
    if ( n_members > 4 ) {
      tstats <- m[which(rownames(m) %in% gs[[1]]),]
      myn <- length(tstats)
      mymean <- mean(tstats)
      mymedian <- median(tstats)
      if ( testtype == "selfcontained" ) { wt <- wilcox.test(tstats) }
      if ( testtype == "competitive" ) { wt <- wilcox.test(tstats,scores) }
      res <- c(name,myn,mymean,mymedian,wt$p.value)
    }
  } , mc.cores = cores)
  res_df <- do.call(rbind, res)
  rownames(res_df) <- res_df[,1]
  res_df <- res_df[,-1]
  colnames(res_df) <- c("n_genes","t_mean","t_median","pval")
  tmp <- apply(res_df,2,as.numeric)
  rownames(tmp) <- rownames(res_df)
  res_df <- tmp
  res_df <- as.data.frame(res_df)
  res_df <- res_df[order(res_df$pval),]
  res_df$logp <- -log10(res_df$pval )
  res_df$fdr <- p.adjust(res_df$pval,method="fdr")
  res_df[order(abs(res_df$pval)),]
  return(res_df)
}

# parametric self-contained
simla <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  dm3 <- runlimma(mval=mval2,design=d,myann=myann)
  dmagg1 <- agg(dm3,cores=4)
  m1 <- dmagg1$gme_res_df[,"mean",drop=FALSE]
  lares1 <- ttenrich(m=m1,genesets=gsets,cores=4,testtype="selfcontained")
  gsig_up3 <- rownames(subset(lares1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(lares1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(lares1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

# parametric competitive
simlac <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  dm3 <- runlimma(mval=mval2,design=d,myann=myann)
  dmagg1 <- agg(dm3,cores=4)
  m1 <- dmagg1$gme_res_df[,"mean",drop=FALSE]
  lares1 <- ttenrich(m=m1,genesets=gsets,cores=4,testtype="competitive")
  gsig_up3 <- rownames(subset(lares1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(lares1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(lares1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

# nonparametric self-contained
simnla <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  dm3 <- runlimma(mval=mval2,design=d,myann=myann)
  dmagg1 <- agg(dm3,cores=4)
  m1 <- dmagg1$gme_res_df[,"mean",drop=FALSE]
  lares1 <- wtenrich(m=m1,genesets=gsets,cores=4,testtype="selfcontained")
  gsig_up3 <- rownames(subset(lares1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(lares1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(lares1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

# nonparametric competitive
simnlac <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  dm3 <- runlimma(mval=mval2,design=d,myann=myann)
  dmagg1 <- agg(dm3,cores=4)
  m1 <- dmagg1$gme_res_df[,"mean",drop=FALSE]
  lares1 <- wtenrich(m=m1,genesets=gsets,cores=4,testtype="competitive")
  gsig_up3 <- rownames(subset(lares1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(lares1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(lares1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

@@

# AL approach parametric self-contained
simal <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  # al pipeline
  medf1 <- chragg(mval=mval2,myann=myann,cores=4)
  magg1 <- agglimma(medf1,d)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  alres1 <- ttenrich(m=m1,genesets=gsets,cores=4,testtype="selfcontained")
  # summarise results
  gsig_up3 <- rownames(subset(alres1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(alres1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(alres1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

# AL approach parametric competitive
simalc <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  # al pipeline
  medf1 <- chragg(mval=mval2,myann=myann,cores=4)
  magg1 <- agglimma(medf1,d)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  alres1 <- ttenrich(m=m1,genesets=gsets,cores=4,testtype="competitive")
  # summarise results
  gsig_up3 <- rownames(subset(alres1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(alres1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(alres1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

# AL approach nonparametric self-contained
simnal <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  # al pipeline
  medf1 <- chragg(mval=mval2,myann=myann,cores=4)
  magg1 <- agglimma(medf1,d)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  alres1 <- wtenrich(m=m1,genesets=gsets,cores=4,testtype="selfcontained")
  # summarise results
  gsig_up3 <- rownames(subset(alres1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(alres1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(alres1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

# AL approach nonparametric competitive
simnalc <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  # al pipeline
  medf1 <- chragg(mval=mval2,myann=myann,cores=4)
  magg1 <- agglimma(medf1,d)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  alres1 <- wtenrich(m=m1,genesets=gsets,cores=4,testtype="competitive")
  # summarise results
  gsig_up3 <- rownames(subset(alres1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(alres1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(alres1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

@@

gres <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- lapply(1:sims,function(i) {
    simgsa(genesetdatabase=gsets, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5,
      frac_probes=0.5, groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  })
  res <- do.call(rbind,res)
  return(res)
})

@@

gres2 <- do.call(rbind,lapply(gres,colMeans))

gres2

@@

### parametric self-contained

@@

lares <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:sims,function(i) {
    simla(genesetdatabase=gsets, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,res)
  return(res)
})

lares2 <- do.call(rbind,lapply(lares,colMeans))

lares2

### parametric competitive

```{r,runlac}

lacres <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:sims,function(i) {
    simlac(genesetdatabase=gsets, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,res)
  return(res)
})

lacres

lacres2 <- do.call(rbind,lapply(lacres,colMeans))

lacres2

### nonparametric self-contained

```{r,runnla}

nlares <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:sims,function(i) {
    simnla(genesetdatabase=gsets, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,res)
  return(res)
})

nlares

nlares2 <- do.call(rbind,lapply(nlares,colMeans))

nlares2

### nonparametric competitive

```{r,runnlac}

nlacres <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:sims,function(i) {
    simnlac(genesetdatabase=gsets, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,res)
  return(res)
})

nlacres

nlacres2 <- do.call(rbind,lapply(nlacres,colMeans))

nlacres2

@@

alres <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:sims,function(i) {
    simal(genesetdatabase=gsets, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,res)
  return(res)
})

@@

alres2 <- do.call(rbind,lapply(alres,colMeans))

alres2

### parametric competitive

```{r,runalc}

alcres <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:sims,function(i) {
    simalc(genesetdatabase=gsets, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,res)
  return(res)
})

alcres

alcres2 <- do.call(rbind,lapply(alcres,colMeans))

alcres2

### nonparametric self-contained

```{r,runnal}

nalres <- lapply(1:nrow(params) , function(j) {
  groupsize = params[j,1]
  delta = params[j,2]
  res <- mclapply(1:sims,function(i) {
    simnal(genesetdatabase=gsets, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
      groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
  },mc.cores=4)
  res <- do.call(rbind,res)
  return(res)
})

nalres

nalres2 <- do.call(rbind,lapply(nalres,colMeans))

nalres2

### nonparametric competitive

```{r,runnalc}
